\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-02-16 arch.tex
"Final draft" on arch.tex:   Addressed Kanna's comments and/or rewrote some sections   The overall flow is here along with the core concepts that needed to be introduced

\item \texttt{step} that accepts new goals requested (if any) on the
  {\em internal} timeline of a reactor and will execute one step of
  deliberation. In the context of \eu for example one deliberation
  step corresponds to a single flaw resolution (Section
  \ref{sec:europa:arch}) while ensuring that the resulting plan is not
  proven inconsistent. This implies that it also includes the
  backtracking of the search in such a case.  The returned value by
  the \texttt{step} function is a boolean which indicates whether the
  reactor needs an extra step to deliberate or if this last step
  resulted in a complete plan within the reactor's look-ahead. On the
  later, the part of the plan that concern this reactor {\em external}
  timelines is provided in order to be dispatched as goals to
  whichever reactor that is reponsible for these. \comment{By
    extending I may have saud things I developed below ... will have
    to check}
\end{itemize}

Both these deliberative functions have different scope and temporal
constraints. The \texttt{synchronize} call is considered as atomic and
will focus only on state inference for the current tick by identifying
the current value of {\em internal} state variables from the latest
updates on the {\em external} state variables. This process propagates
up in the reactor dependency graph at every tick to ensure that all
reactors are aware of the current state of the world. \texttt{step}
relates to a single decision step of the planning process which is
focused on the future evolution of state variables; it can take
multiple steps spanning multiple ticks for a reactor to produce a plan
that is both complete and valid. When such plan is identified
\comment{how? Is there a terminating condition? No more flaws but I
  washoping it would have been defined in \eu section} its {\em external}
timeline(s) are then dispatched to the corresponding reactors which
own these times and which in turn can start their own deliberation
\texttt{step}s to satisfy these new objectives. Consequently execution
occurs as these goals and plans propagate down in the reactor
hierarchy until they eventually become goals for the lowest level
reactor which -- instead of deliberating -- may just transform this
goal into a command for execution in hardware.

@@

While the two processes are separate, they share the same plan
internal to a reactor; consequently at every synchronization cycle the
planning process is informed of the new world state and its impact on
the plan. Conversely, when synchronization occurs after deliberation
steps, its starting point is the last partial plan resulting from
deliberation. This gives an initial frame to the problemm that helps
-- at least initial -- to target the synchronization problem by
ensuring that the new {\em external} state updates are compatible with
this plan. If so, the synchronization propagates these events in the
plan and can uses it as a help on identifying its {\em internal}
timelines based on this plan. If they are not compatibles, then
synchronization identifies a plan inconsistency  reflecting that the
current plan is not valid anymore. \comment{fpy: I am here and will
  develop further}

\comment{following is not what I intended to express whatsoever
  ... synchroniuzation od not deal with dispatching but execution
  tracking ... it is though helped by the structure of the exisiting
  plan and often identifies execution failures}
 Conversely, when the deliberation process eventually finds a
solution, synchronization is informed about generating the planned
state values \comment{should this read as ``...is informed about
  generated state values..'' instead?} on the {\em external} state
variables. This {\em external} plan defines the set of {\em goals} for
this state variable managed by another {\em reactor}.

@@

This assumption, while helpful to reduce the scope of the planning
problem, becomes problematic for planning within a highly dynamic
environment. In a typical robot this implies that the only point where
the planner can be informed about the current world state is prior to
the commencement of a planning cycle by giving to the planner the
initial state and current objectives. During the planning, it is
assumed that the agent can maintin the world as perceived by the
planner as stable. \cite{lemai04, lemai-chenevier2004} attempts to
reduce the impact of planning by allowing local plan repair in the
existing plan. In the eventuality that such repair is not feasible
they require that their robot comes to a full stop until (re-)planning
is complete. \comment{I may need more examples/refs here}. As Fig.
\ref{fig:tick-exec} shows, this assumption is problematic for our
architecture. Synchronization that tracks state evolution can occur
several times during the reactor deliberation process. By making
synchronization and deliberation share the same plan we provide a
solution that relaxes this static world assumption. \comment{Need to
  talk about IDEA and the fact that they had similar aspect even
  though they were never very clear nor well articulated in their
  papers}

@@

{\em Synchronization} and the dispatching of new {\em goals} from
\rx can occur between two steps and alter the plan. These alterations
exogenous to the deliberation process are dynamically impacting the set
of flaws that need to be resolved. Obviously new flaws are created
when a new goal is received from \rx and conversely often the recall
of an objective will remove the corresponding flaws\footnote{although
  if the flaws of this gaol were previously resolved, the removal of
  this goal may create new flaws}. The impact of {\em synchronization}
can be more subtle as its alteration of the plan at the execution
frontier is much more drastic. We can still give a non exhaustive list
of possible alterations of the flaws directly related to {\em
  synchronization}:
\begin{enumerate}
\item Previous flaws directly related to the execution frontier have
  been resolved by the {\em synchronization} solver.
\item The {\em synchronization} solver may have created new flaws that
  were outside of its scope but are now part of the {\em deliberation}
  solver scope.
\item Last observations proved the current plan to be invalid
  requiring the {\em synchronization} solver to relax the existing
  plan in order to identify its current {\em external} state.
\end{enumerate}
Finally as {\em synchronization} occurs when the {\em
  execution frontier} advances, it implies that the reactor planning
scope is further in the future. This need to be reflected by the {\em
  deliberation} solver which need a new deliberation scope one tick
further in the future.

>>>>

\item \texttt{step} that accepts new goals requested (if any) on the
  {\em internal} timeline of a reactor and will execute one step of
  deliberation. In the context of \eu for example one deliberation
  step corresponds to a single flaw resolution (Section
  \ref{sec:europa:arch}) while ensuring that the resulting plan is not
  proven inconsistent. This implies that it also includes the
  backtracking of the search in such a case.  The returned value by
  the \texttt{step} function is a boolean which indicates whether the
  reactor needs an extra step to deliberate or if this last step
  resulted in a complete plan within the reactor's look-ahead. On the
  later, the part of the plan that concern this reactor {\em external}
  timelines is provided in order to be dispatched as goals to
  whichever reactor that is reponsible for these. % \comment{By
    % extending I may have saud things I developed below ... will have
    % to check}
\end{itemize}

Both these deliberative functions have different scope and temporal
constraints. The \texttt{synchronize} call is considered as atomic and
will focus only on state inference for the current tick by identifying
the current value of {\em internal} state variables from the latest
updates on the {\em external} state variables. This process propagates
up in the reactor dependency graph at every tick to ensure that all
reactors are aware of the current state of the world. \texttt{step}
relates to a single decision step of the planning process which is
focused on the future evolution of state variables; it can take
multiple steps spanning multiple ticks for a reactor to produce a plan
that is both complete and valid. How the reactor identifies that its
plan is complete and valid is its responsibility and depend on the way
such reactor resolve its plan during each {\em steps}. The general
idea is that the reactor was able to find a valid strategy tin order
to satisfy as much of its {\em internal} goals as possible. In the
context of \eu, which we are going to develop, this means that the
solver do not identify any more flaw to be resolved in the plan.

 When such plan is identified, its {\em external}
timeline(s) are then dispatched to the corresponding reactors which
own these times and which in turn can start their own deliberation
\texttt{step}s to satisfy these new objectives. Consequently execution
occurs as these goals and plans propagate down in the reactor
hierarchy until they eventually become goals for the lowest level
reactor which -- instead of deliberating -- may just transform this
goal into a command for execution in hardware.

@@

While the two processes are separate, they share the same plan
internal to a reactor; consequently at every synchronization cycle the
planning process is informed of the new world state and its impact on
the plan. Conversely, when synchronization occurs after deliberation
steps, its starting point is the last partial plan resulting from
deliberation. This gives an initial frame to the problem that helps
-- at least initially -- to target the synchronization problem by
ensuring that the new {\em external} state updates are compatible with
this plan. If so, the synchronization propagates these events in the
plan and can uses it as a help on identifying its {\em internal}
timelines based on this plan. If they are not compatibles, then
{\em synchronization} identifies a plan inconsistency  reflecting that the
current plan is not valid anymore. This results on the retraction of
decisions made during {\em deliberation} steps in order to allow 
{\em synchronization} to identify its {\em internal}  state. Such
relaxation can be viewed as a backtracking on the plan search
triggered exogenously by the world evolution. It is the way our
reactor will identify that its current plan failed to execute and can
trigger further deliberations steps in order to find an alternate 
strategy.

Finally, when the deliberation process did identify a complete
solution plan for its current {\em internal} goals. The part of its
plan that applies to its {\em external} timelines is sent as a 
request to \rx that will handle their dispatching to the corresponding 
reactors. If things go smoothly the {\em synchronization} process will
eventually receive {\em external } observations that confirm the
execution of this plan and will allow the plan execution to progress
inside the reactor (by having the part of the plan observed restricted
to the start time of these observations).

@@

This assumption, while helpful to reduce the scope of the planning
problem, becomes problematic for planning within a highly dynamic
environment. In a typical robot this implies that the only point where
the planner can be informed about the current world state is prior to
the commencement of a planning cycle by giving to the planner the
initial state and current objectives. During the planning, it is
assumed that the agent can maintin the world as perceived by the
planner as stable. \cite{lemai04, lemai-chenevier2004} attempts to
reduce the impact of planning by allowing local plan repair in the
existing plan. In the eventuality that such repair is not feasible
they require that their robot comes to a full stop until (re-)planning
is complete.% \comment{I may need more examples/refs here}. 
As Fig. \ref{fig:tick-exec} shows, this assumption is problematic for our
architecture. Synchronization that tracks state evolution can occur
several times during the reactor deliberation process. By making
synchronization and deliberation share the same plan we provide a
solution that relaxes this static world assumption. According to 
\cite{Dias:2003ua, mus06}, IDEA was sharing the same core idea 
where all the solvers of an agent shared the same plan as a basis of
communication and specifically their {\em Reactive planner} --roughly
similar to our {\em synchronization} solver -- was often the only one 
targeting the execution frontier. Although, none of the papers we
found were detailing how this was handled in the
architecture. Moreover, in \cite{Dias:2003ua} we can read:
\begin{quotation}
  [\textellipsis] as long as the planning horizon over which the
  deliberative planner is working never intersects the current
  execution time, deliberative planning can operate in parallel with
  reactive execution and does not affect the reactivity of the agent. 
\end{quotation}
\comment{fpy: This is my interpretation of this sentence at least .... feel
  free to comment in case you disagree}
Which suggest a non overlap between {\em reactive planner} and {\em
  deliberative planners} horizons. In \rx, and more specifically in
our \eu reactor the planning horizon of {\em synchronization} -- 
limited to the current tick -- is a subset of the planning horizon 
of the deliberation steps -- starting at the current tick and with an
extend based on the latency $\lambda$ and look-ahead $\pi$ of 
the reactor. The way we avoid to create conflict between the two is by
a careful control of the scheduling of these two process by the \rx 
agent.

@@

{\em Synchronization} and the dispatching of new {\em goals} from
\rx can occur between two steps and alter the plan. These alterations
exogenous to the deliberation process are dynamically impacting the set
of flaws that need to be resolved. Obviously new flaws are created
when a new goal is received from \rx and conversely often the recall
of an objective will remove the corresponding flaws\footnote{although
  if the flaws of this goal were previously resolved, the removal of
  this goal may create new flaws}. The impact of {\em synchronization}
can be more subtle as its alteration of the plan at the execution
frontier is much more drastic. We can still give a non exhaustive list
of possible alterations of the flaws directly related to {\em
  synchronization}:
\begin{enumerate}
\item Previous flaws directly related to the execution frontier have
  been resolved by the {\em synchronization} solver.
\item The {\em synchronization} solver may have created new flaws that
  were outside of its scope but are now part of the {\em deliberation}
  solver scope.
\item Last observations proved the current plan to be invalid
  requiring the {\em synchronization} solver to relax the existing
  plan in order to identify its current {\em external} state. (This
  case is developed further on the next section)
\end{enumerate}
Finally as {\em synchronization} occurs when the {\em
  execution frontier} advances, it implies that the reactor planning
scope is further in the future. This need to be reflected by the {\em
  deliberation} solver which need a new deliberation scope one tick
further in the future.
